---
title: "Integrating Human Factors Expertise into Development of Automated Vehicles"
subtitle: "Replication package"
author: "Amna Pir Muhammad, Alessia Knauss, Eric Knauss, Jonas Bärgman"
date: "First revision 2024-04-25. This revision `r Sys.Date()`."
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 1
bibliography: refs.bib
link-citations: true
csl: elsevier-harvard.csl
linkcolor: blue
header-includes:
  - \usepackage{amsmath}
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(cache=TRUE)

# set this to num cores on your computer!
CORES = 4 

library(bayesplot)
library(haven)  
library(bnlearn)  
library(gRain) 
library(ggplot2)  
library(dplyr) 
library(brms)
library(tidyverse)
library(mice) 
library(mcmcplots)
library(rethinking)
library(patchwork)
library(viridis)
library(ggthemes)
theme_set(theme_tufte())

options(brms.backend="cmdstanr")
options(mc.cores = CORES) # set num cores
```

# Introduction
The purpose of this document is to show how the analysis of the experiments was done. With this replication package, anyone can download and rerun the analysis, and check our assumptions.


## Load the data
Here we show how to load the dataset (from the `Data/` directory in the GitHub [repository](https://github.com/amnapir/HF-Placement)), which can then be used in the analysis.

```{r, eval=FALSE}
data <- read_sav("Untitled26.sav")
```
# Recode missing values
data <- data %>%
  mutate(across(starts_with("M210_"), ~if_else(. == -9, NA_real_, .))) %>%
  mutate(across(starts_with("M201_"), ~if_else(. == -9, NA_real_, .)))

